\bf On semidefinite-representable sets over valuation fields

### 2024-11-20 article/paper_padic_spec.tex
commit by Simone

Recall that the {\it Newton polygon} $\mathcal{P}$ of a polynomial $P = \sum_{i=0}^{d} a_i T^i \in K[T]$ is
defined as the inferior convex hull of the set $\{(i, \val a_i) \,|\, 0 \le i \le d\}$, {\it i.e.} the
graph of the greatest convex function $\mathcal{P} : [0,d] \to \R$ such that $\mathcal{P}(i) \leq \val a_i$
for $i=1,\ldots,d$. This function is piecewise affine and its slopes are exactly the opposites of the valuations
of the roots of $P$.

@@

\begin{definition}
  \label{def_spectrahedra}
  A \emph{spectrahedron in $K^{d\times d}$} is a set of the form
  $$
  \mathcal{S} = \mathcal{L} \cap K^{d\times d}_+ = \left\{X \in K^{d \times d} : \exists \, x \in K^n, X = A_0+\sum_{x_i A_i}, X \succeq 0\right\}
  $$
  where $\mathcal{L} = A_0+\span{A_1,\ldots,A_n} \subset K^{d \times d}$ is an affine space, for some
  $A_0,A_1,\ldots,A_n \in K^{d\times d}$.
\end{definition}

\begin{remark}
  As in the real case, one often abuses of notation, saying that the preimage
  $S_A = \varphi_A^{-1}(K^{d \times d}_+) = \{x \in K^n : A_0+x_1A_1+\cdots+x_nA_n \succeq 0\}$
  of $K^{d \times d}_+$ under the map $\varphi_A : K^n \to \mathcal{L} \subset K^{d \times d}$, defined as
  $\varphi_A(x) = A_0+x_1A_1+\cdots+x_nA_n$, is a \emph{spectrahedron
  in $K^n$}. Remark that the map $\varphi_A$ is onto $\mathcal{L}$ but not injective,
  unless $A_1,\ldots,A_n$ are linearly indepedent.
\end{remark}

@@

\begin{proposition}
  %%% OLD STATEMENT
  %%% A polyhedron is a spectrahedron.
A polyhedron in $K^n$ is a spectrahedron in $K^n$.
\end{proposition}
\begin{proof}
  Let $P$ be a polyhedron in $K^n$. We show that there exist $d \in \mathbb{N}$ and
  $A_0,A_1,\ldots,A_n \in K^{d \times d}$ such that $P = S_A$. Assume $P = L \cap M$ with
  \begin{equation*}
    \begin{aligned}
    L &= \{x \in K^n : \val \ell_i(x) \geq 0, i=1, \ldots, d\} \\
    M &= \{x \in K^n : \val m_j(x) = +\infty, j=1, \ldots, e\}
    \end{aligned}
  \end{equation*}
  for some $d, e \in \mathbb{N}$. Let $T : K^{\eta} \to K^n$ be a surjective linear map such that
  $\text{im}(T) = M$, with $\eta = \dim(M)$.
  Let $y = (y_1, \ldots, y_\eta)^T \in K^{\eta}$ be variables on $M$, so that $x \in M$ iff
  $x = Ty$ for some $y \in K^{\eta}$. Let $A_0,A_1,\ldots,A_n \in K^{d \times d}$
  be such that
  $$
  A_0+x_1A_1+\cdots+x_nA_n = \diag(\ell_1(x), \ldots, \ell_d(x)).
  $$
  One has that
  $$
  P = L \cap M = \{x \in K^n : x = Ty \in M, A_0+\sum_{i=1}^n x_iA_i \succeq 0\}.
  $$
%%% OLD PROOF
%  Let $\ell_i,m_j \in K[x]_1$, for $i=1,\ldots,d$ and $j=1,\ldots,e$, and let $P = Q \cap L \subset K^n$
%  be a polyhedron as in \Cref{def_polyhedra}, for $Q = \{x \in K^n : \val(\ell_i) \geq 0, i=1,\ldots,d\}$
%  and $L = \{x \in K^n : \val(m_j) = +\infty, j=1,\ldots,e\}$.
%  First, denote by $A(x) = \ell_1 \oplus \cdots \oplus \ell_d \in K[x]^{d\times d}_1$,
%  so that $A(x) = A_0+x_1A_1+\cdots+x_nA_n$ for some (diagonal) $A_i \in K^{d\times d}$.
%  We might assume that $A_0,A_1,\ldots,A_n$ are affinely independent, so that the mapping
%  $\varphi(x) = A(x)$ is injective. 
%  First, remark that $Q=S_A$ and thus $\varphi(Q) = \mathcal{L} \cap K^{d\times d}$, with
%  $\mathcal{L} = A_0+\span{A_1,\ldots,A_n}$ is a spectrahedron.
%  Next, remark that $L \subset K^n$ is the affine space defined by equations $m_1=\ldots=m_e=0$,
%  hence $\varphi(L) \subset K^{d\times d}$ is an affine space. Since $\varphi$ is injective,
%  one has $\varphi(P) = \varphi(Q \cap L) = \varphi(Q) \cap \varphi(L)$ which is the intersection
%  of a spectrahedron with an affine space, thus a spectrahedron.
%  %Define $B = m_1 \oplus \cdots \oplus m_e \in K[x]^{e\times e}_1$
\end{proof}

>>>>

Recall that the \emph{Newton polygon} $\mathcal{P}$ of a polynomial $P = \sum_{i=0}^{d} a_i T^i \in K[T]$ is
defined as the inferior convex hull of the set $\{(i, \val a_i) \,|\, 0 \le i \le d\}$, {\it i.e.} the
graph of the greatest convex function $\mathcal{P} : [0,d] \to \R$ such that $\mathcal{P}(i) \leq \val a_i$
for $i=1,\ldots,d$. This function is piecewise affine and its slopes are exactly the opposites of the valuations
of the roots of $P$.

@@

\begin{definition}
  \label{def_spectrahedra}
  A \emph{spectrahedron in $K^{d\times d}$} is a set of the form
  $$
  \mathcal{S} = \mathcal{L} \cap K^{d\times d}_+ = \left\{X \in K^{d \times d} :
  \exists \, x \in K^n, X = A_0 + \sum_{i=1}^n x_i A_i, X \succeq 0\right\}
  $$
  where $\mathcal{L} = A_0+\span{A_1,\ldots,A_n} \subset K^{d \times d}$ is an affine space, for some
  $A_0,A_1,\ldots,A_n \in K^{d\times d}$.
\end{definition}

\begin{remark}
  As in the real case, one often abuses of notation, saying that the preimage
  $S_A = \varphi_A^{-1}(K^{d \times d}_+) = \{x \in K^n : A_0+x_1A_1+\cdots+x_nA_n \succeq 0\}$
  of $K^{d \times d}_+$ under the map $\varphi_A : K^n \to \mathcal{L} \subset K^{d \times d}$, defined as
  $\varphi_A(x) = A_0+x_1A_1+\cdots+x_nA_n$, is a \emph{spectrahedron
  in $K^n$}. Remark that the map $\varphi_A$ is onto $\mathcal{L}$ but not injective,
  unless $A_1,\ldots,A_n$ are linearly independent.
\end{remark}

@@

\begin{proposition}
  %%% OLD STATEMENT
  %%% A polyhedron is a spectrahedron.
A polyhedron in $K^n$ is a spectrahedron in $K^n$.
\end{proposition}
\begin{proof}
  Let $P = L \cap M$ be a polyhedron in $K^n$, with
  \begin{equation*}
    \begin{aligned}
    L &= \{x \in K^n : \val \ell_i(x) \geq 0, i=1, \ldots, d\} \\
    M &= \{x \in K^n : \val m_j(x) = +\infty, j=1, \ldots, e\}
    \end{aligned}
  \end{equation*}
  for some $d, e \in \mathbb{N}$.
  %Let $T : K^{\eta} \to K^n$ be a surjective linear map such that
  %$\text{im}(T) = M$, with $\eta = \dim(M)$.
  %Let $y = (y_1, \ldots, y_\eta)^T \in K^{\eta}$ be variables on $M$, so that $x \in M$ iff
  %$x = Ty$ for some $y \in K^{\eta}$.
  Let $A_0,A_1,\ldots,A_n \in K^{d \times d}$ be such that
  $$
  \varphi_A(x) := A_0+x_1A_1+\cdots+x_nA_n = \diag(\ell_1(x), \ldots, \ell_d(x)).
  $$
  One has $L = \varphi_A^{-1}(K^{d \times d}_+)$.
  %and hence
  %$$
  %P = \{x \in K^n : x = Ty \in M, \varphi_A(x) \succeq 0\}.
  %$$

  \vspace{1cm}
  In other words, we showed that there exist $d \in \mathbb{N}$ and
  $A_0,A_1,\ldots,A_n \in K^{d \times d}$ such that $P = S_A$, as stated.
%%% OLD PROOF
%  Let $\ell_i,m_j \in K[x]_1$, for $i=1,\ldots,d$ and $j=1,\ldots,e$, and let $P = Q \cap L \subset K^n$
%  be a polyhedron as in \Cref{def_polyhedra}, for $Q = \{x \in K^n : \val(\ell_i) \geq 0, i=1,\ldots,d\}$
%  and $L = \{x \in K^n : \val(m_j) = +\infty, j=1,\ldots,e\}$.
%  First, denote by $A(x) = \ell_1 \oplus \cdots \oplus \ell_d \in K[x]^{d\times d}_1$,
%  so that $A(x) = A_0+x_1A_1+\cdots+x_nA_n$ for some (diagonal) $A_i \in K^{d\times d}$.
%  We might assume that $A_0,A_1,\ldots,A_n$ are affinely independent, so that the mapping
%  $\varphi(x) = A(x)$ is injective. 
%  First, remark that $Q=S_A$ and thus $\varphi(Q) = \mathcal{L} \cap K^{d\times d}$, with
%  $\mathcal{L} = A_0+\span{A_1,\ldots,A_n}$ is a spectrahedron.
%  Next, remark that $L \subset K^n$ is the affine space defined by equations $m_1=\ldots=m_e=0$,
%  hence $\varphi(L) \subset K^{d\times d}$ is an affine space. Since $\varphi$ is injective,
%  one has $\varphi(P) = \varphi(Q \cap L) = \varphi(Q) \cap \varphi(L)$ which is the intersection
%  of a spectrahedron with an affine space, thus a spectrahedron.
%  %Define $B = m_1 \oplus \cdots \oplus m_e \in K[x]^{e\times e}_1$
\end{proof}
